---
title: "Flexdashboard layouts, part 3"  
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  flexdashboard::flex_dashboard:
    orientation: column
    vertical_layout: fill
    
---

```{r echo = FALSE}
# Course: 5210 Communicating Data
# Purpose: Illustrate flexdashboard package layouts, part 3
# Date: August 5, 2019
# Author: Gareth Green

```

```{r echo = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

```

```{r echo = FALSE}
# Load packages
library(flexdashboard)

```


Page 1 - Many dashboards are only one page
=======================================================================

Column {data-width=600}
-----------------------------------------------------------------------

### Chart A - the "---" create a new area on the dashboard

```{r}

```

Output for Chart A would go here

Column {data-width=400}
-----------------------------------------------------------------------

### Chart B - notice how this chart is in a new column

+ If instead used "orientation: row" in YAML this would be below previous chart area  

    - not depend on word "Column" in code, depends on "------"
    - column or row is set in YAML

```{r}

```

### Chart C - notice this chart is to the right because used fill: row

```{r}

```

Output for Chart C would go here


Page 2
=======================================================================

Column {.tabset data-width=250}
-----------------------------------------------------------------------
### Area 1, Tab 1

{.tabset} - creates tab on same page  

+ Multiple tabs are not appropriate in many settings  

```{r}

```

### Tab 2
Stuff on tab 2

```{r}

```

### Tab 3
Stuff on tab 3

```{r}

```

Column {data-width=250}
-----------------------------------------------------------------------
### Area 4

```{r}

```

Output for area 4 would go here


Column {data-width=250}
-----------------------------------------------------------------------
### Area 5

```{r}

```

Output for area 5 would go here


Column {data-width=250}
-----------------------------------------------------------------------
### Area 6

```{r}

```

Output for area 6 would go here


Page 3
=======================================================================

Sidebar {.sidebar}
-----------------------------------------------------------------------

Sidebar for shiny inputs {.sidebar}


Column {data-width=250}
-----------------------------------------------------------------------
### Area 7

```{r}

```

Output for area 7 would go here


Column {data-width=250}
-----------------------------------------------------------------------
### Area 8

```{r}

```

Output for area 8 would go here
