---
title: "week_07_notes"
author: "Andrew Nalundasan"
date: "7/31/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# load libraries
library(tidyverse)
library(ggiraph)
library(vtable)
```

# Overview video

+ Evaluating quality of visualizations and dashboards

    - Best tool to have is your own gut
        - Look at it, and ask if it's easy to interpret
        - What do you think it means?
    - Communication is always geared towards the audience. Not yourself
    - Get rid of legends
        - at minimum: legend.position to move the legend from the side to on the graph
        - label the lines, bars, points instead of relying on legends
    - Dashboards
        - Make it extremely apparent what each graph should show
        - Put the conclusion of the graph as the title
        - use scales package to format numbers
            - "scale", "mark", "label" arguments to format the numbers even more
        - use ggiraph over plotly package for interactive dashboards and graphs

+ Don't just create something and be done with it

    - Must know how to communicate it to the audience
    - Audience wants to know the insight that comes from the data, not about the work that you did
    - "Does the visual get the point across?"
        - Goal: create something that people can interpret and learn from

# Dissecting Visuals

+ Finding the story in the data isn't enough. Also need to communicate the story

+ Use what I want the audience to know to guide my design choices

    1. Type of graph
    2. context to provide
    3. Emphasize and de-emphasize
    4. Order and alignment of information
    5. Titles, labels, and annotations
    
+ When making a visual, consider:

    1. Who is your audience and what do you want them to learn/do?
    2. Choose the type of graph that best highlights your data for your purpose
        - 1 variable: histogram or density
        - summaries of data: bar graph
        - showing a relationship: scatter plot or variation
        - motion: line graph
    3. Order of the visual
        - where are my eyes drawn
        - order the data
        - color, size
        what to emphasize/de-emphasize
    4. Use text, labels, legends, annotations, context appropriately
        - tell the audience what I want them to know
        
+ mean: removes the influence of total number of observations
    
    - removes skewedness of the data
    - makes relative comparisons. normalizes the groups on the same scale
    - relative comparisons show **effectiveness** of differe A/B groups
    
+ relative comparison
    
# Lessons in Storytelling

+ the hard work comes from creating the story out of the data

    - Finding the story is done in the EDA and is the easy part
    - The art comes from finding out how to make a communication
    - it's not a mystery on how to do it. 
    - the mystery is taking the time to learn the audience and give them what they need to know
    
+ storytelling techniques to engage your audience

    1. Move from "showing the data" to "using the data to address an issue"
        - showing data does not provide solutions to the problem
    2. Think about storytelling in movies and plays
        - **Beginning** - introduce characters, relationships, and issues to draw audience in
        - **Middle** - examing alternatives for characters to resolve an issue or conflict
        - **End** - recommend a course of action
    3. Relate to storytelling with data by comparing a movie in my presentation
        - Use the 3-act structure as a model for communication/presentation
    4. Conventional presentation is generally bulleted facts
        - Just the facts. Not linked to context or woven together
        - Showing the data-not telling the story of the data
        - Easier because not finding the story
    5. Every data set has a story
        - FInding and telling a story is more challenging than conventional rhetoric
        - Story tells how and why business practice is changing
            - Resolution/Recommendation: bring business back into balance 
    6. Story can be revealed by asking key questions
        - Main character should usually be the **AUDIENCE**
        - Use stories to engage the audience in a way that "just the facts" cannot
        
    
+ storytelling

    - Beginning:
        - Why important/care?
        - set up
        - context
        - staekholders
        - big picture
        - introduce issue
    - Middle:
        - conflict
        - determine alternatives
        - learn about alternatives
        - consider options
        - weight trade-offs
        - determine recommendation
    - End: 
        - Resolution
        - REcommendation
        - Course of action
        - Restore balance
        - Justified by analysis in middle
        - Commit, but accept input

+ story design, writing

    1. Brain dump first - then fill in and edit to complete thoughts - IDEAS
    2. Determine main topic of each idea into a paragraph
        - Group similar topics and reorder
    3. Review each paragraph
        - Make it flow
    4. Review each sentence
    
+ story telling tactics

    1. Backward design/reverse storyboarding
        - Terentino their ass!
    2. Horizontal logic
        - Ensure paragraphs are in order
    3. Vertical logic
        - Ensure all information in each slide belongs
        - Sentences are in order
    4. Know your audience
        - Speak in the same language as the audience
    5. Fresh perspective
        - Have a 3rd party review and give feedback
        
+ Presentation vs. Visuals

    - Presentation > Dashboard > individual visual
    
+ **LEAVE ENOUGH TIME IN WORK FLOW TO CREATE**

    - Visuals that are:
        - Intentional
        - Meaningful
        - Clear/concise
        - Elegant
    - **AUDIENCE ONLY SEES THE COMMUNICATION**

# Dashboards - Best Practices

+ Dashboard: 
    
    - Collection of visuals that are related to answering a specific problem
    - Contain information to achieve an objective
    - Consolidated on a single page or screen
    - Must consider audience for best design

+ Don't want to try to make a dashboard that shows everything. It won't work

+ Dashboard mistakes

    1. Exceeding the boundaries of a single screen
    2. Supplying inadequate context for the data
    3. Display excessive detail
    4. Expressing measures indirectly is a mistake
    5. Choosing inappropriate display
    6. Introducing meaningless variety
    7. Using poorly designed display
    8. Encoding quantitative information inaccurately
        - Treat each element as an individual visual
    9. Arranging information poorly
    10. Highlighting important information ineffectively or not at all

# Dashboards - Examples

+ using "W7-4-1_bad_dash.RMD" and "W7-4-2_bad_dash.Rmd"

+ dashboard needs to convey a message. 

+ dashboard needs to drive what the audience is meant to learn from it

+ error bars not appropriate for non-technical audience

+ flexdashboard helps create the structure to build dashboards on

# Dashboards - Basic Flexdashboard Layout

+ dashboards should be 1 pagers - no tabs

+ all about the YAML to drive the structure

    - orientation: row vs. column

# Dashboards - Starbucks Storyboard

+ using a dashboard to make a slideshow - do this for QP2?

+ YAML:
    
    - storyboard: true
    - source: embed

+ choropleth map is colored by measure

# Dashboards - Create a basic Flexdashboard



